---
title: "STAT 4207/5207: Midterm Exam"
author: "Test Student"
date: "Fall 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk\$set(echo = TRUE)
# Load core course package libraries
library(tidyverse)
library(MASS)
library(cluster)
```

## Question 1: FBI Dataset Exploration

We load the FBI crime statistics dataset using the system's standard relative data pathway and extract its sample covariance parameters.

```{r question-1}
# Read the dataset securely from the mapped container folder
fbi_data <- read.csv("data/fbi.csv")

# Isolate numeric columns to compute the sample covariance matrix
# (Assumes columns 2 through 5 are numeric for this test case)
numeric_fbi <- fbi_data %>% select(where(is.numeric))
fbi_cov_matrix <- cov(numeric_fbi)

# View covariance dimensions
dim(fbi_cov_matrix)
```

## Question 2: Men's Track Records PCA

We apply Principal Component Analysis (PCA) to the track metrics to explore variance dimensions, making sure to explicitly enable standard feature standardization scaling.

```{r question-2}
# Load the track dataset 
track_data <- read.csv("data/records.men.csv")

# Isolate numeric metrics for dimensional reduction
numeric_track <- track_data %>% select(where(is.numeric))

# Run PCA with mandatory standardization scaling flags active
track_pca <- prcomp(numeric_track, scale. = TRUE)

# Display variance breakdown
summary(track_pca)
```

## Question 3: Wine Quality Cluster Analysis

We implement a partitioning cluster algorithm using K-Means to separate the chemical characteristics of wine into 3 distinct categories.

```{r question-3}
# Load the wine profiling matrix
wine_data <- read.csv("data/winequality.csv")

# Isolate features for partitioning
numeric_wine <- wine_data %>% select(where(is.numeric))

# Enforce numerical convergence reproducibility boundaries
set.seed(42)

# Execute the K-Means algorithm targeting 3 cluster partitions
wine_clusters <- kmeans(numeric_wine, centers = 3, nstart = 20)

# Display within-cluster summary metrics
wine_clusters\$tot.withinss
```
