---
title: "STAT 4207/5207 — Lab 1"
author: "Student group"
output:
  html_document:
    toc: true
    self_contained: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Submission instructions

Keep `Lab1.Rmd` and every supplied data file in the same folder. Complete
the R code and answer the written questions below. Knit the complete report to
HTML (recommended) or PDF on your own computer to check that the code, tables,
plots, and written responses render correctly. Upload only `Lab1.Rmd` to
Gradescope; do not upload the HTML or PDF file. The autograder evaluates the
named R objects, and a teaching assistant grades the written responses in the
submitted `Lab1.Rmd`.

## Written responses — 2 points, graded by a TA

Answer each question below in complete sentences and refer to your R output.
These responses are graded manually by a teaching assistant.

### Exercise 2 — Exploring the tips data

1. How many cases are in the `tips` data? Explain which output shows this.

   **Response:** Replace this text with your answer.

2. How many statistical variables are in the data? Explain why `obs` is an
   observation identifier rather than a statistical variable.

   **Response:** Replace this text with your answer.

3. What types of variables are present? Identify the quantitative and
   categorical variables and explain how the output from `str()`, `summary()`,
   `names()`, or `levels()` supports your classification.

   **Response:** Replace this text with your answer.

4. On which days of the week were tips collected? Cite the relevant factor
   levels or summary output.

   **Response:** Replace this text with your answer.

### Exercise 3 — Scatterplot interpretation

5. Describe what the scatterplot reveals about the relationship between the
   amount of the tip and the total bill. Discuss direction, form, spread, and
   any unusual observations.

   **Response:** Replace this text with your answer.

### Exercise 4 — Simple linear regression

6. Report the intercept and slope of the least-squares regression line and
   interpret both coefficients in the context of total bill and tip. Also
   explain the estimated average tip for a $45 bill and the meaning of the
   slope percentage.

   **Response:** Replace this text with your answer.

## Autograded R work — 8 points

```{r analysis}
# Group members:
# Replace the example names below. This object is recorded for the instructor
# but is not assigned points by the autograder.
group_members <- c("Name 1", "Name 2")

# Keep this file and tips.csv in the same folder while you work.
# Submit this file to Gradescope with the filename Lab1.Rmd.
# Do not rename any requested object.

# -----------------------------------------------------------------------------
# Exercise 1: R as a calculator and R objects (1.6 points)
# -----------------------------------------------------------------------------

# Choose any finite number as your starting value.
start_value <- NULL

# Add 2 to start_value.
after_add <- NULL

# Multiply after_add by 3.
after_multiply <- NULL

# Subtract 6 from after_multiply.
after_subtract <- NULL

# Divide after_subtract by 3. The result should equal start_value.
final_value <- NULL

# -----------------------------------------------------------------------------
# Load the tips data (0.8 point)
# -----------------------------------------------------------------------------

# Read tips.csv as a data frame. Use stringsAsFactors = TRUE so the categorical
# columns are stored as factors, as in the original lab.
tips <- NULL

# -----------------------------------------------------------------------------
# Exercise 2: Getting started with data (2 points)
# -----------------------------------------------------------------------------

# Begin Exercise 2 by using all six functions to explore the data.
dim(tips)
head(tips)
str(tips)
summary(tips)
names(tips)
levels(tips$day)

# Store the number of cases (rows) in n_cases.
n_cases <- NULL

# Store the number of variables in n_variables. Do not count obs, which is only
# an observation identifier.
n_variables <- NULL

# Store the names of the quantitative variables, excluding obs, in a character
# vector. Order does not matter.
quantitative_variables <- NULL

# Store the names of the categorical variables in a character vector.
# Order does not matter.
categorical_variables <- NULL

# Store the days on which tips were collected in a character vector.
# Abbreviations such as "Thu" or full names such as "Thursday" are accepted.
collection_days <- NULL

# -----------------------------------------------------------------------------
# Exercise 3: Scatterplot and interpretation (1.6 points)
# -----------------------------------------------------------------------------

# Create a ggplot2 scatterplot with total bill on the x-axis and tip on the
# y-axis. Store the plot object in tips_plot. Use ggplot() with geom_point(). Do not run install.packages() in this file.
tips_plot <- NULL

# Choose exactly one listed response for each interpretation object.
association_direction <- NULL  # "negative", "none", or "positive"
relationship_form <- NULL       # "approximately linear", "strongly curved", or "no clear form"
spread_pattern <- NULL           # "decreases", "approximately constant", or "increases"

# -----------------------------------------------------------------------------
# Exercise 4: Simple linear regression (1.6 points)
# -----------------------------------------------------------------------------

# Fit tip as the response and totbill as the explanatory variable.
tip_model <- NULL

# Extract the estimated intercept and slope from tip_model.
intercept <- NULL
slope <- NULL

# Use the fitted model to estimate the average tip for a $45 total bill.
predicted_tip_45 <- NULL

# Express the slope as a percentage of the total bill (for example, 10.5 rather
# than 0.105). Store that number in slope_percent.
slope_percent <- NULL
```
